Appendix A

### 2020-09-08 appendixA/appendixA.Rmd
updates

---
title: "Appendix A: Model fitting and selection for tracking data"
output:
  html_document:
      toc: true
bibliography: "../paper/mybibfile.bib"
---

@@

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

@@

fam_dens_plot <- df %>%
  ggplot(., aes(x = mpm)) +
  #geom_line(stat = "density", color = "red", lwd = 1) +
  geom_line(aes(x = mpm, group = fam_g), stat = "density", alpha = 0.4) +
  geom_vline(xintercept = 0, color = "grey") +
  geom_hline(yintercept = 0, color = "grey") +
  ylim(0, 0.05) +
  labs(title = "Family") +
  theme_classic()

@@

# Population vs Individual levels
## Population level
We fit the probability distributions to all the data, so it is population level. We compare fits visually with a qqplot. It is evident with this plot that the lognormal distribution is problematic at the tails and shows a significant overestimation. All the other distributions underestimate the real values, but it seems like the Weibull is closer to the diagonal.

```{r poplevel, eval = FALSE}

pop <- lapply(dist.used, function(x){fitdist(df$mpm, distr = x)})
qqcomp(pop, plotstyle = "ggplot") +
  coord_equal() +
  geom_point(shape = 16, alpha = 0.5)

@@

>>>>

---
title: "Appendix A"
subtitle: "Model fitting and selection for tracking data"
output:
  html_document:
      toc: true
bibliography: "../paper/mybibfile.bib"
---

# Overall description

@@

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE, results = 'hide')
```

@@

fam_dens_plot <- df %>%
  filter(., fam_g != "unknown") %>% 
  ggplot(., aes(x = mpm)) +
  #geom_line(stat = "density", color = "red", lwd = 1) +
  geom_line(aes(x = mpm, group = fam_g), stat = "density", alpha = 0.4) +
  geom_vline(xintercept = 0, color = "grey") +
  geom_hline(yintercept = 0, color = "grey") +
  ylim(0, 0.05) +
  labs(title = "Family") +
  theme_classic()

@@

```
We have chosen five common probability distributions, which are readily used in animal-mediated seed dispersal simulations. These are an exponential distribution [@jones_closing_2017], a gamma distribution, Weibull distribution [@morales_effects_2006], and lognormal distribution [@levey_effects_2005]. The goal of this project is to evaluate variation in movement and how that translates to differences in seed dispersal estimates. Our first comparisson is made between population level and individual level estimates. Not all individual birds had information on family group, therefore, for our second comparisson between population level and family level fits, we used a reduced dataset with only data from individual for which we had family group information. We assessed fits visually through QQ plots.

# Population and Individual level comparissons

At the population level analysis, we observe that the lognormal distribution has considerable overestimation of distances towards the tail. Although this is also worth noting in the QQ plots for individuals, not all of the individual have an overestimation, and for some other distributions seem to do a better job at characterizing the pattern.

```{r poplevel, fig.height = 3}

pop <- lapply(dist.used, function(x){fitdist(df$mpm, distr = x)})

qqpop <- qqcomp(pop, plotstyle = "ggplot")

qqpop +
  coord_equal() +
  geom_point(shape = 16, alpha = 0.5) +
  labs(title = "Population level QQ plot")
  scale_color_viridis_d()

@@

```{r}
inds <- unique(as.character(df$Bird_ID))
#inds

for(i in 1:length(inds)){
  fit.df <- df %>% 
    dplyr::filter(., Bird_ID == inds[i])

  fit <- lapply(dist.used, function(x){fitdist(fit.df$mpm, distr = x)})

  assign(paste(inds[i]), fit)

  fit.prms <- build.fits.df(fit) %>% 
  mutate(param = param,
         dist = dist,
         data = inds[i])

  prms.df <- bind_rows(prms.df, fit.prms)
}
```

```{r fig.height=5}

inds.models.data <- NULL
ind.models <- list(B1, B13, B19, B22, B28, B29, B3, B30, B49, B5, B7, B84)
chr.models <- c("B1", "B13", "B19", "B22", "B28", "B29", "B3", "B30", "B49", "B5", "B7", "B84")
for(i in 1:12){
  x <- ind.models[[i]]

  qqdata <- qqcomp(x, plotstyle = "ggplot")$data %>% 
    mutate(data = chr.models[i])
  names(qqdata) <- c("theoretical", "fdist", "empirical", "individual")
  inds.models.data <- bind_rows(inds.models.data, qqdata)
}

inds.models.data%>% 
  ggplot(., aes(x = theoretical, y = empirical, color = fdist)) +
  facet_wrap(~individual)+
    geom_point(cex = 2, alpha = 0.6) +
    geom_abline(intercept = 0, slope = 1) +
    labs(x = "Model Quantiles",
         y = "Empirical Quantiles") +
  scale_color_viridis_d() + 
  theme(legend.position = "top") +
  guides(color = guide_legend(title = "Distribution")) +
  coord_equal()

## Information criteria assessment of individual level distribution fits
Need to add the description here of how we calculated these information criteria and the reasoning as to why we decided to go with just the BIC, because of Jose's paper. Explain here what the mixed distribution model means, vs the others that use one distribution but allow variation in the parameters.

```{r}
all_indiv_prms <- prms.df %>% 
  mutate(kpars = ifelse(dist == "exp", 1, 2),
         BIC = log(n)*kpars - 2*loglik,
         AIC = 2*kpars - 2*loglik,
         AICc = AIC + (2* kpars * (kpars+1)/(n-kpars-1))) %>% 
  group_by(data) %>% 
  mutate(deltaBIC = signif(BIC - min(BIC), 3),
         deltaAIC = signif(AIC - min(AIC), 3),
         deltaAICc = signif(AICc - min(AICc), 3))
#saveRDS(all_indiv_prms, file = "outputs/prms_indiv.RDS")

prms.df %>% 
  dplyr::select(., loglik, n, dist, data) %>% 
  distinct() %>% 
  mutate(kpars = ifelse(dist == "exp", 1, 2),
         BIC = log(n)*kpars - 2*loglik,
         AIC = 2*kpars - 2*loglik,
         AICc = AIC + (2* kpars * (kpars+1)/(n-kpars-1))) %>% 
  group_by(data) %>% 
  mutate(deltaBIC = signif(BIC - min(BIC), 3),
         deltaAIC = signif(AIC - min(AIC), 3),
         deltaAICc = signif(AICc - min(AICc), 3)) -> popind.IC
```
